---
title: "CLIC DCC — Progress Report"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  word_document:
    toc: true
  pdf_document:
    toc: true
    latex_engine: xelatex
  html_document:
    theme: flatly
    toc: true
    toc_float: true
params:
  data: NULL
  scope: "All countries"
  colors: NULL
  study: "Study"
  pages: NULL
  total_studies: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=4)
library(dplyr); library(ggplot2); library(knitr)
d <- params$data
PAL <- params$colors
if (is.null(PAL)) PAL <- c("Complete"="#4CAF50","In Process"="#ffcc33","Incomplete"="#E0413B",
         "N/A"="#1F1F1F","Not started"="#D2D2D2")

SHORT <- c("Total Studies"="CLIC Total Studies","Genomic Pipeline"="Genomic Pipeline",
           "Phenotype Data"="Phenotype Data","Epi Data"="Epi Data",
           "IARC DCC Transfer"="IARC DCC","CoordinatingAnalysisCenter"="Coord. Center")
disp <- function(p) ifelse(p %in% names(SHORT), unname(SHORT[p]), p)

# Tracker study labels: / and & join separate studies (COG AALL0232 & AALL0434 = 2)
split_tracker <- function(x){ x<-as.character(x)
  vapply(x, function(v){ if(is.na(v)||trimws(v)=="") return(0L)
    p <- strsplit(v, "[/&]")[[1]]; p <- p[nzchar(trimws(p))]; length(p) }, integer(1)) }

# Classify each study (unit) into one overall status
study_cat <- function(dd){
  if (is.null(dd) || !nrow(dd)) return(dd[0,])
  dd |> group_by(page, study) |>
    summarise(non_na = sum(status!="N/A"),
              done   = sum(status=="Complete"),
              active = sum(status %in% c("Complete","In Process","Incomplete")),
              .groups="drop") |>
    mutate(units = split_tracker(study),
           cat = case_when(non_na==0 ~ "Complete",
                           done==non_na ~ "Complete",
                           active==0 ~ "Not started",
                           TRUE ~ "In progress"))
}

has_proj <- !is.null(d) && "project" %in% names(d)
has_canc <- !is.null(d) && "cancer"  %in% names(d)
PAGES <- if (!is.null(d) && nrow(d) > 0) unique(as.character(d$page)) else character(0)
gl_studies <- if (has_proj) sum(split_tracker(unique(d$study[d$page=="Genomic Pipeline" & d$project=="GlobALL"]))) else NA
```

**Scope:** `r params$scope`  
**CLIC total studies:** `r if (!is.null(params$total_studies)) params$total_studies else length(unique(d$study))`  **·  Countries:** `r length(unique(d$country))`  
**Overall step progress:** `r round(mean(d$status=="Complete")*100)`%  **·  Not started:** `r round(mean(d$status=="Not started")*100)`%
`r if (has_proj) paste0("**·  GlobALL pipeline studies:** ", gl_studies)`

## Status mix by tab

```{r}
if (!is.null(d) && nrow(d) > 0) {
  dd <- d |> filter(page != "CoordinatingAnalysisCenter")
  lv <- disp(setdiff(PAGES, "CoordinatingAnalysisCenter"))
  dd |> count(page, status) |>
    mutate(page = factor(disp(page), levels = lv)) |>
    ggplot(aes(page, n, fill = status)) +
    geom_col(width=.7) + scale_fill_manual(values = PAL, drop = FALSE) +
    labs(x=NULL, y="Cells", fill=NULL) +
    theme_minimal(base_size=12) + theme(legend.position="top")
} else {
  cat("No data available to display plots.\n")
}
```

## Completion by tab (study units)

Counts are **study units** (a study is counted once; combined entries like "COG AALL0232 & AALL0434" count as two). *Complete* = every milestone in that tab is done for the study; *In progress* = some but not all done; *Not started* = none started. The Coordinating/Analysis Center tab is excluded.

```{r}
if (!is.null(d) && nrow(d) > 0) {
  tabs <- setdiff(PAGES, "CoordinatingAnalysisCenter")
  ps <- study_cat(d[d$page %in% tabs, ])
  base_tab <- ps |> group_by(page) |>
    summarise(Studies=sum(units),
              Complete=sum(units[cat=="Complete"]),
              `In progress`=sum(units[cat=="In progress"]),
              `Not started`=sum(units[cat=="Not started"]), .groups="drop")
  base_tab$Tab <- disp(base_tab$page); base_tab$page <- NULL
  rows <- base_tab[, c("Tab","Studies","Complete","In progress","Not started")]
  if (has_proj && any(d$page=="Genomic Pipeline" & d$project=="GlobALL")) {
    g <- study_cat(d[d$page=="Genomic Pipeline" & d$project=="GlobALL", ])
    grow <- data.frame(Tab="GlobALL Pipeline", Studies=sum(g$units),
      Complete=sum(g$units[g$cat=="Complete"]),
      `In progress`=sum(g$units[g$cat=="In progress"]),
      `Not started`=sum(g$units[g$cat=="Not started"]), check.names=FALSE)
    rows <- rbind(rows, grow)
  }
  ord <- c(); for (p in tabs) { ord <- c(ord, disp(p)); if (p=="Genomic Pipeline") ord <- c(ord, "GlobALL Pipeline") }
  rows$Tab <- factor(rows$Tab, levels=ord); rows <- rows[order(rows$Tab), ]
  kable(rows, row.names = FALSE)
} else {
  cat("No data available to display metrics table.\n")
}
```

## GlobALL pipeline

```{r results='asis'}
if (has_proj && any(d$page=="Genomic Pipeline")) {
  gp  <- d[d$page=="Genomic Pipeline", ]
  gl  <- gp[gp$project=="GlobALL", ]
  tab <- data.frame(
    Population   = c("All genomic studies", "GlobALL project only"),
    Studies      = c(sum(split_tracker(unique(gp$study))), sum(split_tracker(unique(gl$study)))),
    `Step progress` = c(paste0(round(mean(gp$status=="Complete")*100),"%"),
                        if (nrow(gl)) paste0(round(mean(gl$status=="Complete")*100),"%") else "—"),
    check.names  = FALSE)
  print(kable(tab, row.names = FALSE))
} else {
  cat("*No Project column in the current data — GlobALL breakdown unavailable.*\n\n")
}
```

## Cancer types (genomic studies)

```{r results='asis'}
if (has_canc && any(d$page=="Genomic Pipeline")) {
  st <- unique(d[d$page=="Genomic Pipeline", c("study","cancer")])
  types <- trimws(unlist(strsplit(st$cancer[nzchar(st$cancer)], ",")))
  types <- types[nzchar(types)]
  if (length(types)) {
    ct <- as.data.frame(table(`Cancer type`=types), stringsAsFactors=FALSE)
    ct <- ct[order(-ct$Freq), ]; names(ct)[2] <- "Studies"
    print(kable(ct, row.names = FALSE))
  } else cat("*No cancer-type data in the current scope.*\n\n")
} else {
  cat("*No cancer-type data available.*\n\n")
}
```

## Flagged studies (Notes)

```{r}
if (!is.null(d) && nrow(d) > 0) {
  hasinst <- "institution" %in% names(d) && any(nzchar(d$institution))
  flagged <- if (hasinst) d |> filter(notes != "") |> distinct(country, institution, study, notes)
             else d |> filter(notes != "") |> distinct(country, study, notes)
  if (nrow(flagged)) kable(flagged,
     col.names = if (hasinst) c("Country","Institution","Study","Note") else c("Country","Study","Note")) else
    cat("No notes flagged in the current scope.")
} else {
  cat("No data available to scan for flagged notes.\n")
}
```

---

<small>Generated from the CLIC DCC Progress Dashboard. Status reflects whatever has been entered in <code>clic_data.xlsx</code>; blank cells are shown as "Not started" and are awaiting confirmation.</small>
