---
title: "CLIC DCC — Progress Report"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float: true
  word_document:
    toc: true
  pdf_document:
    toc: true
    latex_engine: xelatex
params:
  data: NULL
  scope: "All countries"
  colors: NULL
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE, fig.width=9, fig.height=4)
library(dplyr); library(ggplot2); library(knitr)
d <- params$data
PAL <- params$colors
if (is.null(PAL)) PAL <- c("Complete"="#4CAF50","In Process"="#ffcc33","Incomplete"="#E0413B",
         "N/A"="#1F1F1F","Unknown"="#D2D2D2")
PAGES <- unique(as.character(d$page))
```

**Scope:** `r params$scope`  
**Studies:** `r length(unique(d$study))`  **·  Countries:** `r length(unique(d$country))`  
**Overall complete:** `r round(mean(d$status=="Complete")*100)`%  **·  Still unknown:** `r round(mean(d$status=="Unknown")*100)`%

## Status mix by page

```{r}
d |> count(page, status) |>
  mutate(page = factor(page, levels = PAGES)) |>
  ggplot(aes(page, n, fill = status)) +
  geom_col(width=.7) + scale_fill_manual(values = PAL, drop = FALSE) +
  labs(x=NULL, y="Cells", fill=NULL) +
  theme_minimal(base_size=12) + theme(legend.position="top")
```

## Completion by page

```{r}
d |> group_by(Page = factor(page, levels = PAGES)) |>
  summarise(Complete=sum(status=="Complete"),
            `In process`=sum(status=="In Process"),
            Incomplete=sum(status=="Incomplete"),
            `N/A`=sum(status=="N/A"),
            Unknown=sum(status=="Unknown"),
            `% complete`=round(mean(status=="Complete")*100), .groups="drop") |>
  arrange(Page) |> kable()
```

## Flagged studies (Notes)

```{r}
hasinst <- "institution" %in% names(d) && any(nzchar(d$institution))
flagged <- if (hasinst) d |> filter(notes != "") |> distinct(country, institution, study, notes)
           else d |> filter(notes != "") |> distinct(country, study, notes)
if (nrow(flagged)) kable(flagged,
   col.names = if (hasinst) c("Country","Institution","Study","Note") else c("Country","Study","Note")) else
  cat("No notes flagged in the current scope.")
```

---

<small>Generated from the CLIC DCC Global Progress Dashboard. Status reflects whatever has been entered in <code>clic_data.xlsx</code>; blank/Unknown cells are awaiting confirmation.</small>
